Exponential Family Markets for Information Aggregation

### 2014-02-12 WriteUp/maxentpredmkts.tex
added liquidity proof

\section{Budget-limited Aggregation} \label{sec:budgets}

In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. We assume that the traders trade in multiple instances of the market. As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution \sindhu{I think we need this for Lemma \ref{lem:damage}}. We measure the error in prediction using the standard log loss.

@@

>>>>

\section{Budget-limited Aggregation} \label{sec:budgets}
In this section, we consider the evolution of the market state when traders are budget-limited and their budgets are known to the market maker. We consider a cost-function based market with cost function $C$ and payoff $\phi(x)$ with non-negative payoff. We assume that the traders trade in multiple instances of the market. As before, the market price is interpreted as a probability density over the outcome space by interpreting the share vector as the natural parameter of an exponential family distribution. We measure the error in prediction using the standard log loss.

@@

\subsection{Cost Function with a Liquidity Parameter}
We may introduce a liquidity parameter to the cost function as $C_{\lambda}(\theta)=\frac{1}{\lambda} C(\lambda\theta)$. We would like to find an expression for a lower bound on expected payoff. We would also like to find an expression for final market state for exponential utility traders with exponential family beliefs.
\subsection{lower bound on expected payoff}
The prediction market is set up with cost function $C_{\lambda}(\theta)$  and payoff for a unit share vector as $\phi(x)$. To find a lower bound on expected payoff, we first note that
\begin{eqnarray*}
\E_{x\sim p(x;\beta)}[C_{\lambda}(\theta)-\theta\phi(x)]&=&C_{\lambda}(\theta)-\theta\nabla T(\beta)\\
&=&\frac{1}{\lambda} T(\theta\lambda)-\theta\nabla T(\beta)\\
&=&\frac{1}{\lambda} [T(\theta\lambda)-\theta\lambda\nabla T(\beta)-T(\beta)+\beta\nabla T(\beta)]+\frac{1}{\lambda} T(\beta)-\frac{\beta}{\lambda}\nabla T(\beta)\\
&=&\frac{1}{\lambda} D_{T}(\theta\lambda,\beta)+\frac{1}{\lambda} [T(\beta)-\beta\nabla T(\beta)]
\end{eqnarray*}¥ 
where $D_{T}(x_{1},x_{2}) = T(x_{1})-T(x_{2})-\nabla T(x_{2})(x_{1}-x_{2})$ is the Bregman Divergence between $x_{1}$ and $x_{2}$ based on $T$.

Thus, the expected payoff  for a trader with exponential family belief parameter $\beta$ who moves the market state from $\theta$ to $\theta'$ can be written as
\begin{eqnarray*}
\E_{x\sim p(x;\beta)}\left[C_{\lambda}(\theta)-\theta\phi(x)-(C_{\lambda}(\theta')-\theta'\phi(x))\right]=\frac{1}{\lambda}[D_{T}(\theta\lambda,\beta)-D_{T}(\theta'\lambda,\beta)]
\end{eqnarray*}¥
We have $\theta'=\frac{\lambda}{\lambda+a}(\beta/\lambda)+\frac{a}{\lambda+a}\theta$ where $a$ is the utility parameter
\begin{eqnarray*}
D_{T}(\theta'\lambda,\beta)&\leq& \frac{\lambda}{\lambda+a} D_{T}(\beta,\beta)+\frac{a}{\lambda+a} D_{T}(\theta\lambda,\beta)\\
D_{T}(\theta\lambda,\beta)-D_{T}(\theta'\lambda,\beta)&\geq& \frac{\lambda}{a} D_{T}(\theta'\lambda,\beta)
\end{eqnarray*}¥

Thus, the expected payoff is at least $\frac{1}{a} D_{T}(\theta'\lambda,\beta)$.
